---
title: "Statistical Inference Project Part 1"
author: "Jacob Townson"
date: "July 20, 2015"
output: pdf_document
---

```{r include=FALSE}
library(knitr)
library(ggplot2)
```


#Overview

In this project I will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Note, I did not make an appendix section because I merged my code with the explanations, and still did not go over 3 pages. 

# Part 1

**Show the sample mean and compare it to the theoretical mean of the distribution.**

Well, we know that the theoretical mean is $\frac{1}{\lambda}$. Well, since $\lambda = .2$, the theoretical mean is $5$. Now I just need the simulation to compare it to. In order to do our simulation, I will run this code:

```{r}
set.seed(2200)
mns = NULL
for (i in 1 : 1000) mns = c(mns, mean(rexp(40, .2)))
```

Note in the above code I set the seed that I will be working with. This is in order to make sure my simulations don't vary from code chunk to code chunk.

Now, if I take the mean of the means, it can be compared to the theoretical mean of $5$:

```{r}
mean(mns)
```

One can see that the mean of all of the simulated means is $5.00045$. This is extremely close to our theoretical mean.

# Part 2

**Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.**

We know that the theoretical variance is $\frac{1}{\lambda^2}$. Just like the theoretical mean, because $\lambda = .2$, so we know that the theoretical variance is going to be $25$, and the theoretical standard deviation to be $5$. To do the simulation, I will use a similar method as I did for the mean. 

```{r}
set.seed(2200)
vars = NULL
for (i in 1 : 1000) vars = c(vars, var(rexp(40, .2)))
mean(vars)
```

After taking the mean of all of the variances, we can see that the samples all have around a variance of $25$ and standard deviation of $5$. Thus our sample and theoretical are extremely close once again.

# Part 3

**Show that the distribution is approximately normal.**

First, let's take a look at a simple histogram of $1000$ random exponentials with $\lambda = .2$. 

```{r}
set.seed(2200)
hist(rexp(1000, .2))
```

This doesn't look very normal at all. However, if we take the mean of a large number of simulations using the code below, we find something different.

```{r}
set.seed(2200)
mns = NULL
for (i in 1 : 1000) mns = c(mns, mean(rexp(40, .2)))
hist(mns)
```

In this histogram, we get something that looks much more normal centered at our theoretical mean of $5$. So we can see through these examples how the Central Limit Theorem works for an exponential distribution.










